Generating Datasets about Building Designs \\Based on a Multilayer Building Ontology

### 2015-04-30 article-02-main.tex
First version submitted

All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (instead of \name{IFCXXX} there should be the name of IFC version, e.g. \name{IFC4\_ADD1}):

@@

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. An attribute is defined with a data\-type, cardinalities and can be marked as optional. An entity data\-type should be converted to an \name{owl:\-Class} which is subclass of its superclass (if any) or \name{expr:\-Entity\-Class}. It was decided to make all entity properties be functional (the maximum cardinality is 1).

Entity attribute cardinalities can be only from 0 to 1, even if this attribute is a set of values. In this case the attribute type is an external aggregated data\-type, as in:

%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcProperty a owl:Class ; rdfs:subClassOf expr:EntityClass .
ifc:IfcComplexProperty a owl:Class ; rdfs:subClassOf ifc:IfcProperty ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:allValuesFrom ifc:LIST_1_3_IfcLengthMeasure ] ;
    # only for Extended layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:coordinates;
                      owl:cardinality "1"^^xsd:integer ] .

# only for Standard & Extended layers
[] a owl:AllDisjointClasses ;
    owl:members ( ifc:IfcComplexProperty ifc:IfcSimpleProperty ) .
\end{lstlisting}

@@

Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass} which are subclasses of \name{expr:\-Collec\-tion\-Class}. The latter has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}. Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}. Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:

% In order to declare these types, new classes were created in a similar way as Ordered List Ontoloty \cite{olo:ontology}:
% \begin{itemize}
% \item Class \name{expr:\-Collec\-tion\-Class} has four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index} and a nonfunctional property \name{expr:slot} with type \name{expr:Slot}.
% \item Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}.
% \item Classes \name{expr:SetClass}, \name{expr:ListClass}, \name{expr:ArrayClass} and \name{expr:BagClass} are subclasses of \name{expr:\-Collec\-tion\-Class}.
% \end{itemize}

% Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:

\begin{lstlisting}
# all layers:
 ifc:SET_2_UNDEFINED_OF_IfcProfileDef a owl:Class ;
    rdfs:subClassOf ifc:SET_2_UNDEFINED , ifc:SET_OF_IfcProfileDef .
 ifc:SET_2_UNDEFINED, SET_OF_IfcProfileDef rdfs:subClassOf expr:SetClass .

# only for layer Extended:
 ifc:SET_2_UNDEFINED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .
 ifc:SET_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:allValuesFrom ifc:SLOT_OF_IfcProfileDef ].
 ifc:SLOT_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ; owl:onProperty expr:item ; owl:allValuesFrom ifc:IfcProfileDef ] .
\end{lstlisting}

@@

\subsubsection{Header Section} IFC data is exchanged as \emph{STEP physical files} (SPF). SPF format is defined in ISO 10303-21 \cite{ISO10303-21}. A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription} \name{FileName} and \name{FileSchema}. %described in \cite{buildingSMART:ifc-header}. 
They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:

@@

\subsubsection{Data Section} In a SPF data section each entity is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types have a GUID in IFC. They are converted into URI resources with format \name{GUID\_\textless{}guid\textgreater{}} where \name{guid} is in the hexadecimal string representation (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\textless{}lineNumber\textgreater{}} (in a user defined namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:

\begin{lstlisting}
:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C
    a               ifc:IfcPropertySet ; # entity type
    ifc:globalId    [ a ifc:IfcGloballyUniqueId ; # defined type
                     rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
    ifc:hasProperties  _:b11 .
_:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ; # entity type
    expr:size "3"^^xsd:integer ;
    expr:itemType ifc:IfcProperty ;
    expr:slot _:b12, _:b13, _:b14 .
_:b12 a ifc:SLOT_OF_IfcProperty ;
    expr:item [ a ifc:IfcPropertySingleValue ; # entity type
                ifc:name "initial_GUID"^^xsd:string ;
                ifc:nominalValue  _:b15 ] .
\end{lstlisting}

@@

\vspace{2mm}
\begin{tabular}{|l|l|}
\hline
Ontologies: & \texttt{http://drumbeat.cs.hut.fi/owl/} \\
\hline
Example datasets & \\
- IFC files: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
- RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
\hline
Converter & \\
- service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
- open source: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
\hline
\end{tabular}

\section{Summary and future work}
\label{sec:conclusion}

We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
derived from the IFC schema with the \texfbf{IFC2LD} converter that can also produce RDF datasets from IFC files.

The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project. We are exploring the possibilities to use the IFC2LD converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.

>>>>

All constructs related to EXPRESS, STEP or IFC specifications belong to separate namespaces shown below with prefixes \name{expr:}, \name{step:} and \name{ifc:} accordingly (\name{IFCXXX} should be replaced by the name of an IFC version, e.g. \name{IFC4\_ADD1}):

@@

@@

Any aggregated type in \ifcowl{} is a subclass of \name{expr:SetClass},
\name{expr:ListClass}, \name{expr:ArrayClass} or \name{expr:BagClass}. Their common superclass \name{expr:\-Collec\-tion\-Class} has one nonfunctional property \name{expr:slot} with type \name{expr:Slot} and four functional properties: \name{expr:isOrdered}, \name{expr:item\-Type}, \name{expr:\-size}, \name{expr:\-start\-Index}, \name{expr:\-end\-Index}. Class \name{expr:Slot} has two functional properties \name{expr:item} (the actual collection item value), \name{expr:index} and also optional functional properties \name{expr:previous}, \name{expr:next}. Below is a fragment of definition of type \name{SET [2:?] OF IfcProfileDef}:

\begin{lstlisting}
# all layers:
 ifc:SET_2_UNLIMITED_OF_IfcProfileDef a owl:Class ;
    rdfs:subClassOf ifc:SET_2_UNLIMITED , ifc:SET_OF_IfcProfileDef .
 ifc:SET_2_UNLIMITED, SET_OF_IfcProfileDef rdfs:subClassOf expr:SetClass .

# only for layer Extended:
 ifc:SET_2_UNLIMITED rdfs:subClassOf rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:minCardinality "2"^^xsd:integer ] .
 ifc:SET_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ;
    owl:onProperty expr:slot ; owl:allValuesFrom ifc:SLOT_OF_IfcProfileDef ].
 ifc:SLOT_OF_IfcProfileDef rdfs:subClassOf [ a owl:Restriction ; owl:onProperty expr:item ; owl:allValuesFrom ifc:IfcProfileDef ] .
\end{lstlisting}

\subsubsection{Entity Datatypes} correspond to classes in object-oriented languages. Each type may have one super\-type. An attribute is defined with a data\-type, can be a list or a set with \emph{any} cardinalities and can be marked as optional. An entity data\-type should be converted to a subclass of its super\-class (\name{expr:\-Entity\-Class} is the most generic). It was decided to make maximum cardinality of any entity property be \name{1}. In case if the attribute type is a list or a set then a new external aggregated data\-type is created. Below is an example for a type with attributes \name{Profiles : SET [2:?] OF IfcProfileDef} and \name{Label : OPTIONAL IfcLabel}:

%\begin{lstlisting}[caption={Entity data\-types}, label=lst:ifcOWL-defined-types]
\begin{lstlisting}
ifc:IfcCompositeProfileDef a owl:Class ; rdfs:subClassOf ifc:IfcProfileDef ;
  # only for Standard, Extended layers
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:allValuesFrom ifc:SET_2_UNLIMITED_IfcProfileDef ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:allValuesFrom ifc:Label ] ;
  # only for Extended layer
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:profiles;
                      owl:cardinality "1"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:minCardinality "0"^^xsd:integer ] ;
    rdfs:subClassOf [ a owl:Restriction; owl:onProperty ifc:label;
                      owl:maxCardinality "1"^^xsd:integer ] .
\end{lstlisting}

@@

\subsubsection{Header Section.} IFC data is exchanged as \emph{STEP physical files} (SPF) format of which is defined in ISO 10303-21 \cite{ISO10303-21}. A SPF header section includes a list of STEP-specific metadata entities with types \name{FileDescription}, \name{FileName} and \name{FileSchema}. %described in \cite{buildingSMART:ifc-header}. 
They are exported as objects of STEP entities, some attributes are duplicated as Dublin Core (DC) metadata annotations, for example:

@@

\subsubsection{Data Section.} In a SPF data section each entity is identified with a line number.
%, an entity data\-type and a list of values. 
All entities of \name{Ifc\-Root}-derived types have a GUID in IFC. They are converted into URI resources with format \name{GUID\_\emph{\textless{}guid\textgreater{}}} where \emph{guid} is in the hexadecimal string representation (not including the character \name{\$}). Regarding to anonymous IFC entities, it can be chosen whether to export them as blank nodes or create them a URI in a format \name{LINE\_\emph{\textless{}lineNumber\textgreater{}}} (in a user defined namespace). Note that line numbers are used only for linking entities inside a SPF, and they are unstable. Naming is required when the dataset is used together with OWL 2 EL/QL as these profiles do not allow anonymous individuals. Below is an example of IFC data exported to an RDF dataset compatible with OWL 2 RL:

\begin{lstlisting}
:GUID_1916794F-5605-4499-9AB0-F155DE8D3B6C     # entity as an URI
    a            ifc:IfcPropertySet ;          # entity type
    ifc:globalId [ a ifc:IfcGloballyUniqueId ; # property with defined type
                   rdf:value "0P5dbFLWL4cPgmyLNUZJji"^^xsd:string ] ;
    ifc:hasProperties  _:b11 .                 # property w/ aggregated type
_:b11 a ifc:SET_1_UNLIMITED_OF_IfcProperty ;
    expr:size "3"^^xsd:integer ;
    expr:itemType ifc:IfcProperty ;
    expr:slot _:b12, _:b13, _:b14 .
_:b12 a ifc:SLOT_OF_IfcProperty ;
    expr:item [ a ifc:IfcPropertySingleValue ;  # entity as a blank node
                ifc:name "initial_GUID"^^xsd:string ;
                ifc:nominalValue  _:b15 ] .
\end{lstlisting}

@@

% \vspace{0.5\topsep}
\noindent
\begin{scriptsize}
\begin{center}
    \def\arraystretch{1.2}
    \setlength{\tabcolsep}{6pt}
    \begin{tabularx}{0.90\textwidth}{|l|X|}
        \hline
        Ontologies: & \texttt{http://drumbeat.cs.hut.fi/owl/} \\
        \hline
        Example files & \\
        - IFC data: & \texttt{http://drumbeat.cs.hut.fi/ifc/} \\
        - RDF datasets: & \texttt{http://drumbeat.cs.hut.fi/rdf/} \\
        \hline
        IFC2LD Converter & \\
        - online service: & \texttt{http://drumbeat.cs.hut.fi/home/} \\
        - download binary: & \texttt{http://drumbeat.cs.hut.fi/ifc2ld/} \\
        - source code: & \texttt{https://github.com/Web-of-Building-Data/Ifc2Rdf} \\
        \hline
    \end{tabularx}
\end{center}
\end{scriptsize}

\section{Summary and future works}
\label{sec:summary}

We have presented \ifcowl{}, a multilayer ontology of building designs based on the widely adopted IFC standard. The ontology is 
derived from the IFC schema with the \ifcld{} converter that can also produce RDF datasets from IFC files.

The \ifcowl{} ontologies and \ifcrdf{} datasets are currently used in a variety of Linked Data oriented use cases \cite{torma2013semantic} developed in the DRUMBEAT project. We are exploring the possibilities to use the \ifcld{} converter in the generation of metadata for \ifcrdf{} datasets and, in some cases, also linksets between datasets.
